---
title: "George_Smith_IST719_Lab2"
author: "George Smith"
date: "10/17/2021"
output: pdf_document
---
# Author George Smith 
# Purpose: Lab 2, data interogration or data exploration 

# Data exploration Lab 2
# this code only works for interactive files as this file is on my PC it is not relevant 

# install package
```{r}
#install.packages("vioplot")
#install.packages("plotrix")
```

```{r}
#fname <- file.choose()
```
# supplemented the above code with the following line 
```{r}
tips <- read.csv("C:/Users/GeorgeSmith/Documents/Syracuse/IST 719/tips.csv"
                 , header = TRUE
                #, StringASFactors = FALSE
                )
```
# view the name of columns
```{r}  
colnames(tips)
```
# lets you edit data in spreadsheet view  
```{r}
fix(tips)
```

# lets you view data cant edit 
```{r}
#view(tips)
```

# structure of the data 
```{r}
str(tips)
```

# call rows and columns 
```{r}
tips[1,]
tips[,1]
tips[3,3]
tips[1:3,]
```
#tells us how many variables are in this subset 
```{r}
length(tips [1:3, 2])
```

# number of dimensions in a data set 
```{r}
dim(tips)
```
# subsets the dim function 
```{r}
dim(tips)[1]
```

#call a column 
```{r}
tips$time 
```
# differnt way to call the columns 
```{r}
tips[,"time"]
```
# plot the total bull column
```{r}
plot(tips$total_bill)
```
# sorts the data 
```{r}
plot(sort(tips$total_bill))
```
# boxplot 
```{r}
boxplot(tips$total_bill)
```
# histogram 
```{r}
hist(tips$total_bill)
```

# density plot
```{r}
d <- density(tips$total_bill)
plot(d)
polygon(d, col = "orange")
```
# boxplot of tips for asynch 
```{r}
boxplot(tips$tip)
```
# have multiple chars on the same plot 
```{r}
par(mfrow = c(2,2))
boxplot(tips$total_bill)
hist(tips$total_bill)
d <- density(tips$total_bill)
plot(d)
polygon(d, col = "orange")
```
# add a vioplot
```{r}
library(vioplot)
```

```{r}
vioplot(tips$total_bill)
```

```{r}
unique(tips$sex)
```
#subet for just males 
```{r}
tips.M <- tips[tips$sex == "Male", ]
```
# subet only females 
```{r}
tips.F <- tips[tips$sex == "Female", ]
```

```{r}
par(mfrow = c(2,1), mar = c(2,3,1,2))
boxplot(tips.F$tip, horizontal = T, ylim = c(1,10))
boxplot(tips.M$tip, horizontal = T, ylim = c(1,10))
```

# working with JSON files 
```{r}
fname <- "C:/Users/GeorgeSmith/Documents/Syracuse/IST 719/tweet.formated.json"
```
# import library 
```{r}
library(jsonlite)
```
# read in data 
```{r}
raw.tweet<- fromJSON(fname, flatten = FALSE)
```
# analyze data 
```{r}
str(raw.tweet)
```

```{r}
names(raw.tweet)
```

```{r}
raw.tweet$text
```

```{r}
raw.tweet$user$followers_count
```

```{r}
raw.tweet[["user"]]
raw.tweet[["user"]]$followers_count
raw.tweet[["user"]][["follower_count"]]
```

# read in new file 
# based on error message appears to be issues with the underlying file 
```{r}
fname3 <- "C:/Users/GeorgeSmith/Documents/Syracuse/IST 719/tweets5814.json"
#con <- file(fname3, open ="r")
#tweets <- stream_in(con)
#close(con)
```

```{r}
#dim(tweets)
```

```{r}
#tweets$text [1:3]
```

```{r}
#boxplot(log10(tweets$user$followers_count), horizontal = TRUE ) 
```
# distribution work
```{r}
task.time <- c(rnorm(n=30, mean=30, sd =2.25)
               ,rnorm (n= 30, mean =25, sd = 1.5))
hist(task.time)
```

```{r}
status <- c(rep("AMA", 30), rep("PRO", 30))
df <- data.frame(time = task.time, status= status)

```

```{r}
df.grouped <- aggregate(df$time, list(df$status), mean)
```

```{r}
colnames(df.grouped) <- c("stat", "time")
```

```{r}
df.grouped
```

```{r}
barplot(df.grouped$time, names.arg = df.grouped$stat)
```

```{r}
M.grouped <- tapply(df$time, list(df$status), mean)
class(M.grouped)
```

# returns a matrix or vector same as aggregate 
```{r}
tapply(df$time, list(df$status), range)
```

```{r}
summary(task.time)
```

```{r}
range(task.time)
```

```{r}
aggregate(df$time, list(df$status), summary)
```

```{r}
table(df$status)
table(round(df$time,2))
```

```{r}
df$sex <- sample(c("M","F"), 60, replace = T)

```

```{r}
aggregate(df$time, list(df$status, df$sex), mean)
M <- tapply(df$time, list(df$sex, df$status), mean)
M <- tapply(df$time, list(df$sex, df$status), mean)
barplot(M, beside = TRUE)
```

# reshaping data with tidyr 
```{r}
library(tidyr)
```

```{r}
n <- 5 
year <- 2001:(2000 + n)
q1 <- runif(n = n, min = 100, max = 120)
q2 <- runif(n=n, min = 103, max = 130)
q3 <- runif(n=n, min = 104, max = 140)
q4 <- runif(n=n, min = 108, max = 150)

df.wide<- data.frame(year, q1,q2,q3,q4)
gather(df.wide, qt, sales, q1:q4)
```

```{r}
df.wide %>% gather(qt, sales, q1:q4)
```

```{r}
df.long <- df.wide %>% gather(qt, sales, q1:q4)
o <- order(df.long$year, df.long$qt)
df.long <- df.long[o,]
```

```{r}
df <- data.frame (cat=rep(c("tap", "reg", "zed", "vum"),3)
                  , group = rep(letters[7:9], 4)
                  , x= 1:12)
```

```{r}
spread(df,cat,x)
```

# using rect function to build a custom plot 
```{r}
library(plotrix)
n <- 70
age.min <- 1
age.max <- 90
age.range <- c(age.min, age.max)
m <- round(rescale(rbeta(n,5,2.5), age.range), 0)
hist(m)
f <- round(rescale(rbeta(n,5,2.0), age.range), 0)
x <- age.min : age.max
f.y <- m.y <- rep (0, length(x))
```

```{r}
m.tab <- table(m)
m.y[as.numeric((names(m.tab)))] <- as.numeric(m.tab)

f.tab <- table(f)
f.y[as.numeric((names(f.tab)))] <- as.numeric (f.tab)

age.freqs <- data.frame (ages = x, males =m.y, females =f.y)

max.x <- round(1.2 * max(age.freqs[,2:3]), 0)
plot(c(-max.x, max.x), c(0,100), type = "n", bty = "n", xaxt = "n" 
     , ylab = "age", xlab = "freq", main = "sample age distribution")

grid()
last.y <- 0
for (i in 1:90) {
    rect(xleft = 0, ybottom = last.y, xright = -age.freqs$males[i]
       , ytop =age.freqs$ages[i], col = "lightblue2", border = NA)
       
    rect(xleft = 0, ybottom = last.y, xright = -age.freqs$females[i]
       , ytop =age.freqs$ages[i], col = "lightpink", border = NA)
    
    last.y <- age.freqs$ages[i]
}
```

```