---
output:
  pdf_document: default
  html_document: default
---
# Initial data analysis
```{r}
myCars <- mtcars
myCars
```
# Step 1: what is the highest hp 
## 1. What is the highest hp 
```{r}
  max(myCars$hp)
```
## 2. Which car has the highest hp 

```{r}
myCars[myCars$hp == max(myCars$hp),]
```

# Step 2: Explore mpg 
## 3. What is the highest mpg? 
```{r}
max(myCars$mpg)
```    

## 4. Which car has the highest mpg 
```{r}
myCars[myCars$mpg ==max(myCars$mpg),]
```

## 5. Create a sorted dataframe, based on mpg
```{r}
sortedmpg <- myCars[order(myCars$mpg),]
sortedmpg
```

# Step 3: Which car has the "best" combination of mpg and hp
## 6. What logic did you Use?
### I decided I would define efficeny as the best combination of mpg/hp.
```{r}
myCars$efficency <- myCars$mpg/myCars$hp
```

## 7. Which car?
```{r}
myCars[myCars$efficency ==max(myCars$efficency),]
```

# Step 4: Which car has the best car combination of mpg and hp, where mpg and hp must be given equal weight? 
```{r} 
  scale(myCars$hp)
  scale(myCars$mpg)
  myCars$combo <- scale(myCars$hp)+scale(myCars$mpg)
  myCars[myCars$combo ==max(myCars$combo),]
```
  
  