---
output:
  pdf_document: default
  html_document: default
---
---
title: "George_Smith_HW8"
author: "George Smith"
date: "6/7/2021"
output: pdf_document


### 1. Read in the data 
### 2. 2. You	have	the	option	of	saving	the	file	save	this	file	to	your	computer	and	read	it	into	R,	or reading the	data	directly	from	the	web	into	a	data	frame.

Antelope <- readXL("C:/Users/GeorgeSmith/Desktop/Antelope.xlsx", 
  rownames=FALSE, header=TRUE, na="", sheet="Antelope", stringsAsFactors=TRUE)

### 3. You	should	inspect	the	data	using	the	str()	command	to	make	sure	that	all	of	the	cases	have	been	read	in	(n=8	years	of observations)	and	that	there	are	fourvariables.	
```{r}
str(Antelope)
```
### 4. Create	bivariate	plots	of	number	of	baby	fawns versus	adult	antelope	population,	the	precipitation	that	year,	and	the	severity	of	the	winter.	Your	code	should	produce	three	separate	plots.	Make	sure	the	Y-axis	and	X-axis	are	labeled.	Keeping	in	mind	that	the	number	of	fawns is	the	outcome	(or	dependent)	variable,	which	axis	should	it	go	on	in	your	plots?
```{r}
plot(Antelope$AdultPop, Antelope$FawnCount)
```

```{r}
plot(Antelope$Precip, Antelope$FawnCount)
```

```{r}
plot(Antelope$WinterSeverity, Antelope$FawnCount)
```

### 5. Next,	create	three	regression	models	of	increasing	complexity	using	lm().	In	the	first	model,	predict	the	number	of	fawns	from	the	severity	of	the	winter.	In	the	second	model,	predict	the	number	of	fawns from	two	variables	(one	should	be	the	severity	of	the	winter).	In	the	third	model	predict	the	number	of	fawns from	the	three	other	variables.	Which	model	works	best?	Which	of	the	predictors	are	statistically	significant	in	each	model? If	you	wanted	to	create	the	most	parsimonious	model	(i.e.,the	one	that	did	the	best	job	with	the	fewest	predictors),	what	would	it	contain?

```{r}
model1 <- lm(formula=FawnCount ~ WinterSeverity, data=Antelope)
  summary(model1)
```

## winter severity has a p value of .036 which is less than .05 so it would be considered significant. the r squared is .47 
```{r}
model2 <- lm(formula=FawnCount ~ WinterSeverity + Precip, data=Antelope)
  summary(model2)
```

## winter severity in not conisdered significant in this model as the p value is .19. Precip would be considered significant as the p value is .008. this model has an r squared value of .86
```{r}
model3 <- lm(formula=FawnCount ~ AdultPop + Precip + WinterSeverity, data=Antelope)
  summary(model3)
```
## all input variables in this model are considered significant as each p value is less than .05 the r squared value is .96


## The most parsimonious model has 1 input variable Precip. The r-squared value of .82 means this 1 value explains 82 percent of the fawn count 

```{r}
model4 <- lm(formula=FawnCount ~ Precip, data=Antelope)
  summary(model4)
```

## example predictions 

```{r}
test=data.frame(Precip=9)
predict(model4,test,type="response")
```

```{r}
test=data.frame(Precip=16)
predict(model4,test,type="response")
```
