---
title: "George_Smith_HW6"
author: "George Smith"
date: "5/16/2021"
output: pdf_document
---
### install 
#install.packages("ggplot2")
#library("ggplot2")

###step 1: Load the Data 

```{r}
air <- airquality 
```

###step 2: Clean the data 

```{r}
remove_na <- function(df, n=0){
  df[rowSums(is.na(df)) <= n,]
}

air2 <- remove_na(air)
summary(air2)
```

### Step 3: Understand the data distribution
## Ozone Histogram 
```{r}
Ozone <- ggplot(air2,aes(x=Ozone))
Ozone <- Ozone + geom_histogram(binwidth=10,color="blue",fill="green",na.rm=TRUE)
Ozone
```
## Solar.R Histogram 
```{r}
Solar <- ggplot(air2,aes(x=Solar.R))
Solar <- Solar + geom_histogram(binwidth=10,color="black",fill="yellow",na.rm=TRUE)
Solar
```

## Wind Historgram 

```{r}
Wind <- ggplot(air2,aes(x=Wind))
Wind <- Wind + geom_histogram(binwidth=1,color="red",fill="blue",na.rm=TRUE)
Wind
```

## Temp Histogram
```{r}
Temp <- ggplot(air2,aes(x=Temp))
Temp <- Temp + geom_histogram(binwidth=1,color="black",fill="red",na.rm=TRUE)
Temp
```
## Ozone boxplot  
```{r}
OzoneBP <- ggplot(air2,aes(x=Ozone, y=Month))
OzoneBP <- OzoneBP + geom_boxplot(na.rm=TRUE)
OzoneBP
```
## wind values boxplot 
```{r}
wind <- factor(air2$Wind)
windBP <- ggplot(air2, aes(x=wind, y=Month))
windBP <- windBP + geom_boxplot(na.rm=TRUE)
windBP
```
### Explore how the data changes over time 
## create data column 

```{r}
date <- as.Date(with(air2, paste(1973, air2$Month, air2$Day,sep="-")), "%Y-%m-%d")
```
## Ozone date chart 
```{r}
Ozone2 <- ggplot(air2, aes(x=reorder(Ozone,date),y=date,group=1))
Ozone2 <- Ozone2 + geom_line(color="green",na.rm=TRUE)+geom_point()
Ozone2 <- Ozone2 + theme(axis.text.x=element_text(angle=90,hjust=1))
Ozone2
```
## temp date chart 
```{r}
temp2 <- ggplot(air2, aes(x=reorder(Temp,date),y=date,group=1))
temp2 <- temp2 + geom_line(color="orange",na.rm=TRUE)+geom_point()
temp2 <- temp2 + theme(axis.text.x=element_text(angle=90,hjust=1))
temp2
```

## Wind date chart 
```{r}
wind2 <- ggplot(air2, aes(x=reorder(Wind,date),y=date,group=1))
wind2 <- wind2 + geom_line(color="purple",na.rm=TRUE)+geom_point()
wind2 <- wind2 + theme(axis.text.x=element_text(angle=90,hjust=1))
wind2
```

## Solar date chart 

```{r}
solar2 <- ggplot(air2, aes(x=reorder(Solar.R,date),y=date,group=1))
solar2 <- solar2 + geom_line(color="red",na.rm=TRUE)+geom_point()
solar2 <- solar2 + theme(axis.text.x=element_text(angle=90,hjust=1))
solar2
```

## one chart with 4 lines 
```{r}
plot <- ggplot(air2,aes(x=date),group=1)
plot <- plot + geom_line(aes(y=Ozone,color="Ozone"),na.rm=TRUE)
plot <- plot + geom_line(aes(y=Temp,color="Temp"),na.rm=TRUE)
plot <- plot + geom_line(aes(y=Wind,color="Wind"),na.rm=TRUE)
plot <- plot + geom_line(aes(y=Solar.R,color="Solar.R"),na.rm=TRUE)
plot
```

## Step 4: Look at all the data via Heatmap
```{r}
heat<- ggplot(air2, aes(x=Day,group=Day)) 
heat<- heat + geom_tile(aes(y = Ozone), color="red",na.rm=TRUE)  
heat<- heat + geom_tile(aes(y = Temp), color="green",na.rm=TRUE)
heat<- heat + geom_tile(aes(y = Wind), color="purple",na.rm=TRUE)
heat<- heat + geom_tile(aes(y = Solar.R), color="orange",na.rm=TRUE)
heat
```

### step 5 Look	at	all	the	data	via	a	scatter	chart
```{r}
scatter	<- ggplot(air2,aes(x=Wind,y=Temp),na.rm=TRUE) 
scatter	<- scatter+geom_point(aes(size=Ozone,color=Solar.R),na.rm=TRUE) 
scatter
```

### Step 6 Final Analysis 

## Do you see any patterns after expoloring the data 
# 
## What was the most useful visualization 
#
