---
title: "George_Smith_HW3"
author: "George Smith"
date: "5/3/2021"
output: pdf_document
---
## Step 1: Create a function (named readStates)to read a CSV file into R
#1. Note that you are to read  URL, not a file local to your computer
#2. The file is a dataset on state populations 

```{r}
readStates <- function(){
  url <- "https://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv"
    states_csv <- read.csv(url)
    return(states_csv)
}
``` 

## Step 2: Clean the dataframe 
# 3. Note the issues that need to be fixed 
# 4. within your function make sure there are 51 rows, Make sure there are only 5 columns with the columns having the following names 
# 5.Make sure the last four columns are numbers 

```{r}
newdf <- readStates()
newdf <- newdf[c(-1:-8, -60:-66), ]
newdf <- newdf[,-6:-10]
colnames(newdf)<- c("stateName","base2010","base2011","Jul2010","Jul2011")

dim(newdf)

newdf$base2010 <- as.numeric((gsub(",", "", newdf$base2010)))
newdf$base2011 <- as.numeric((gsub(",", "", newdf$base2011)))
newdf$Jul2010 <- as.numeric((gsub(",", "",newdf$Jul2010)))
newdf$Jul2011 <- as.numeric((gsub(",", "", newdf$Jul2011)))
``` 
## Step 3: Store and Explore the dataset 
##6. Store the dtatset into a dataframe, called dfStates 
##7 Test your dataframe by calculating the mean for the July2011 data, by doing: mean(dfStates$Jul2011)

```{r}
dfStates <- data.frame(newdf)
mean(dfStates$Jul2011)
```
## Step 4: Find the state with the Highest Population 
## 8. Based on the July2 011 data, what is the population of the state with the hihest population? What is the name of that state?
## 9. sort the data, increasing order, based on the July 2011 data

```{r}
max(dfStates$Jul2011)
which.max(dfStates$Jul2011)
dfStates[5,]
dfStates[order(dfStates$Jul2011,decreasing=FALSE),]
```
##Step 5 Explore the distribution of the states 
## 10. Write a function that takes two parameters. The first is a vector and the second is a number 
## 11. The function will return the percentage of the elements within the vector that is less than the same 
## 12. For	example,if the	vector	had	5	elements	(1,2,3,4,5),	with	2	being	the	number passed	into	the	function,	the	function	would	return	0.2	(since	20%	of	the	numbers	were	below	2).
## 13. Test	the	function	with	the	vector	‘dfStates$Jul2011Num’,	and	the	mean of dfStates$Jul2011Num’.

```{r}
percentage <- function(vec,num)
{ 
  
  flag <- 0 
  
  i <- length(vec) 
  
  {     if(vec[i] < num) {flag <- flag+1}  
    
  } 
  
  return ((flag/length(vec))*100) 
  
}

percentage(dfStates$Jul2011,37253956) 
percentage(mean(dfStates$Jul2011),5000000)
```


